On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2016-01-22 MS.tex
clarify list of outsanding issues, including in the theoretical sections

\input{ms-common-defs}

\date{}

% Todo:
% Ditch abbrvos for SLA, wood density, Hmax?
% Fix plots
%   -- major revision of code -- > lots of repetition, hard to follow
%   -- Figure 2: add * against traits used in meta-analysis: no
% - seach on XXX, footnote

\bibliographystyle{mee}
\mstype{Research Article}
\keywords{maximum height, photosynthetic rate, ontogenetic stage, functional trait, growth strategy, seed mass, specific leaf area, wood density}

@@

\item We propose that traits should influence growth in a way that depends on the size of individual plants. We outline mechanisms and hypotheses based on new theoritical work, and test these predictions in tree species using a meta-analysis of 103 studies ($>$ 500 correlations) for five traits (specific leaf area, wood density, maximum height, seed mass and maximum assimilation rate). We also recorded data for 14 others traits commonly used in the trait literature. To capture the effects of plant size, we tested for a shift in the direction of correlation between growth rates and traits across three ontogenic stages: seedling, sapling and adult.

% Recent theoretical work ? instead of new theory?

\item Results were consistent with predictions, even with some limitations arising from unequal numbers of observation across stages. Specific leaf area was correlated with relative growth rate (RGR) in seedlings but not in adult plants. Correlations of growth with wood density and assimilation rate were not affected by stage. Seed mass and maximum height were correlated with RGR only in one plant stage category: seedlings and adults, respectively. 
\item Despite a sizeable literature on traits and growth, major knowledge gaps across size still exist; for one-third of traits the majority ($>$ 75\%) of correlations with growth come from a single stage. Clearly, there is scope for strengthening our understanding on the link between traits and growth through adding further empirical studies for traits and size classes underrepresented in the literature.

% not sure to keep the end of this sentence

\item Synthesis: We show how trait-growth correlations change in a predictable way with plant size. Our understanding of plant strategies should shift away from attributing species as having slow or fast growth throughout their life, in favour of attributing growth trajectories.

%Clearly, there is scope for strengthening our understanding on the link between traits and growth through adding further empirical studies for traits and size-classes underrepresented in the literature.

\end{enumerate}

\clearpage

@@

Meanwhile the possibility has been raised that trait-growth correlations may change systematically with size \citep{King-1999,Falster:2011ii, Ruger:2012jv, Iida:2014ep, Iida:2014hq}. Size has long been known to play a central role in organismal design and function \citep{Vogel:1988ux, Vogel:2003wb}, including that of trees \citep{Farnsworth:1995im, Givnish:1995ta}. As plants grow larger, a greater amount of the energy is devoted towards building and maintaining non-photosynthetic tissues \citep{Givnish:1995ta}. A recent model by \citet{Falster:2011ii} suggests this change shifts the benefits and costs of some trait-based trade-offs, such that the net effect of the trait on growth changes with size. Thus the absence of a correlation between SLA and growth at larger sizes may in fact be expected, when potential influences of traits on plant function are considered \citep{Falster:2011ii}. While recent empirical studies support the idea that the relationship between traits and growth rates may change with size (\citealt{Iida:2014ep, Iida:2014hq}), the trend and mechanisms behind this variation remain unclear.

In this paper, we investigate for five prominent traits (specific leaf area = SLA, wood or stem tissue density = WD, maximum leaf photosynthetic capacity per unit area = Aarea, asymptotic or maximum height = Hmax, and seed mass):
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions derived from a recently developed mechanistic model \citep{Falster:2011ii} (see ``Theoretical expectations'')?
\end{enumerate}
To address these questions we used a meta-analytic approach, compiling data from diverse studies reporting correlation coefficients between traits and growth. While most studies investigating the link between traits and growth have not primarily been concerned with the effects of plant size, they have generally reported trait-growth correlations for identifiable stage classes (i.e. seedling, sapling, and adult). We therefore used plant stage as a surrogate for plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms. 
We chose to assess our hypotheses via a meta-analysis because there already exist a considerable number of studies with salient information. Using these studies, we were able to assess our hypotheses across a much broader number of species and vegetation types than would have been possible in an experiential study. While, we anticipated a reasonably-sizeable literature on trait and growth to be available, we were uncertain about the representation of all trait-stages combinations. An additional goal of our study was therefore to identify gaps in the knowledge needed to improve our understanding of the link between trait and growth.

@@

Based on the above model and this single assumption about growth, we outline five hypotheses, which arise by considering the influence of trait variation on elements of equations  \ref{eq:eq0}-\ref{eq:eq4} (see Table \ref{tab:trade-offs}) and how these may interact with plant size. These hypotheses reflect a direct mechanistic causation between traits and growth rate. It is possible that other correlations may arise indirectly. Nevertheless, for purposes of this paper we adopt working hypotheses arising only from direct mechanistic causation. With only XXX exception, the predicted effects of the trait are also similar when growth is expressed on relative of absolute terms, and as diameter, height or mass growth (Table \ref{tab:trade-offs} (TRUE XXXX?).

\footnote{The correlation with RGR$_{\rm mass}$ is negative but becomes positive with AGR$_{\rm mass}$. For SLA, WD, Hmax, and Aarea, the direction of the predicted relationship is not affected by the growth measurement used. However RGR and AGR are fundamentally different for a lot of ecologists. So we adopted a two steps approach, first we were conservative by focusing exclusively on RGR, then we compared results with those of a more inclusive dataset with both RGR and AGR. => We have different expectation for SLA in AGRmass for seedling: but none of our data are in AGRmass, they are mainly in RGRmass (Fig. \ref{fig:figA8}). Why a change with WD when it is measured in mass? All WD seedling are measured in RGRmass, and there are some mass data for both other stages too. }

@@

Plants with high SLA can deploy a unit of leaf area more cheaply. However, high-SLA species also have shorter leaf lifespan, and must use more energy replacing leaves lost via turnover \citep{Wright:2004jb}. These two effects occur within the terms  $\frac{\ud B}{\ud t}$ and $\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$ in the above model. Whether plant growth rate increases with SLA depends on how the relative magnitude of the benefit (cheap leaf construction, via $\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$ ) and cost (higher leaf turnover, via $\frac{\ud B}{\ud t}$) of increased SLA change with size (Table \ref{tab:trade-offs}). 
During early life, when leaf comprises a large part of the plant, decreasing SLA has an overwhelmingly positive effect on height and diameter growth rate, because the effect of reducing $\frac{\ud M_{\rm l}}{\ud A_{\rm l}}$ has a large effect compared to the other terms in eq. \ref{eq:daldmt}. For mass growth, cheaper leaf construction (high SLA) translates directly into faster relative growth rate in seedlings (See Appendix \ref{app:growth-equation} eq. \ref{eq:eq1_seedlings_RGR}). As plants increase in size, however, the cost of building sapwood ($\frac{\ud M_{\rm s}}{\ud A_{\rm l}}$) increases disproportionately, decreasing the benefit of cheaper leaf construction in eq.  \ref{eq:daldmt}. Consequently, the relative effect of leaf turnover comes to dominate at larger sizes, and as such, the effect of SLA on height and diameter growth shifts from positive to negative as plants increase in size.

@@

Cheaper stem construction (low WD or stem density) increases the marginal rate of leaf deployment per total mass increment ($\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$), and consequently increases growth rate in leaf area, height, and diameter (eqs. \ref{eq:eq2}-\ref{eq:eq4}; Table \ref{tab:trade-offs}). Disadvantages associated with cheap stem construction take the form of higher mortality and so are not reflected in growth equations. Under this scenario, wood density is negatively correlated to growth rate across all plant stages. Because the relative expenditure on stem compared to leaf increases as plants grow larger, the relationship between WD and growth may also strengthen in larger plants.

@@

As plants approach maximum height they increasingly divert energy away from growth and into reproduction (see \citealt{Thomas:1996do,Thomas-2011, Wenk:2015jz}). This effect occurs in the term $\frac{\ud M_{\rm t}}{\ud B}$ in the above model eqs. (\ref{eq:eq2}-\ref{eq:eq4}). For seedlings, where all individuals are investing solely in growth, we expect no influence of Hmax on growth (i.e. $\frac{\ud M_{\rm t}}{\ud B} =1$). But once some species begin to produce seed, then species closer to their Hmax value are expected to be allocating more to seed production and less to vegetative growth. Hence when comparing across plants at a given size, species with smaller Hmax are expected to show slower vegetative growth rates (Table \ref{tab:trade-offs}).

@@

The main effect of increased seed size is to produce a larger overall seedling. Under similar light conditions, RGR decreases with size because the ratio of leaf area to total biomass decreases along plant size. Thus, larger seedlings are predicted to have a lower RGR during the earliest phases of growth \citep{Turnbull:2012ew}. However, as plants are growing we expect the effect of this initial size to be minimal, and the correlation between seed size and growth rate should disappear. Thus for larger plants, seed mass is not expected to have any distinct influence on growth rates  (Table \ref{tab:trade-offs}). However, the predicted relationship between growth and seed mass in seedlings reverses if AGR$_{\rm mass}$ is used instead of RGR$_{\rm mass}$.

@@

We expect greater photosynthetic capacity to translate into faster growth rates by increasing the rate of biomass production (Table \ref{tab:trade-offs}), at least at high light where Aarea is measured. This effects occurs within the term  $\frac{\ud B}{\ud t} $ in the above model. The effects of benefits (higher assimilation) and costs (higher respiration) of increased Area occur within the same element of eq. \ref{eq:eq0}, and because assimilation is typically higher than respiration in high light, the net effect will be positive (Table \ref{tab:trade-offs}). It is uncertain how much this effect might translate to faster growth in low light. It might greatly depend on the level of light available.

\clearpage

@@

% bib <- bibtex::read.bib("references/metaanalyses.bib")
% paste(sort(names(x$keys)), collapse = ', ')
% These studies dropped because not in list of top traits:
% Bohlman:2006eo, Cochard:2007bn, Coley:1983hk,  Comas:2002wt, Comas:2004jza, King:2006dg, Poorter:2001fj, Schulze:2006ww,  Tjoelker:1999gp, Tobner:2013cv,

\subsubsection*{Plant stage categories}\label{plant-stage-categories}

We used categorical groupings by plant ``stage'' to test our hypotheses regarding the effects of traits on growth rates. The authors of the 103 articles used in this analysis reported their plants as belonging to any of five different stages: ``seedling'', ``juvenile'', ``sapling'', ``adult'', ``mixed''. While these categories have been established according to the expertise of the authors, we were concerned about the risk of assigning plants of similar size to different categories when compiling data across multiple studies and vegetation types. This risk turned out to be small: overall, we found that the original thresholds in size/age used by the authors to distinguish different stages were relatively compatible (Fig. \ref{fig:fig2}). In particular, there was a clear break between seedlings and saplings in both age (at 1 yr) and height (at 0.5 m). There was also a clear break between saplings and adults in diameter (at 10 cm). Comparing these values to the bivariate relationships observed between variables (age, height, diameter) in the ``Biomass And Allometry Database'' \citep{Falster:2015}, demonstrated that the thresholds identified above for each stage transition also aligned relatively well with one another (Fig. \ref{fig:figA1}). Thus we adopted the above values as cut-offs for defining stage boundaries.

Using the identified boundaries among stages we were able to assign all studies into one of three classes: ``seedling'', ``sapling'', or ``adult''. The group ``juvenile'' was largely overlapping with the three others, and so we merged this category with one or the other on a case by case basis. Using the above cut-offs, we were also able to assign a stage category to studies for which no stage was distinctly mentioned by the authors (5\% of the articles, noted as ``mix'' in Fig. \ref{fig:fig2}). Only 7 of 103 articles were ambiguous on their plant stage attribution. Half of these equivocal cases involved the juvenile category. We attributed a study to a new stage category only when 95\% of the recorded range of sizes for that study lay within a particular stage category. Other equivocal cases were caused by ambiguity in the usage of ``adult'' by the authors. Some plants with a large stature were not described as adult by authors because they were still non reproductive \citep{King:2006he}. Conversely plants with small stature -- because stunted by wind for example \citep{Stratton:2001ck} -- were sometimes assessed as adult when reproductive. For purposes of the present paper, we attributed a study to the adult stage when the authors identified the plants as reproductive, and to sapling stage when they identified the plants as non-reproductive, regardless of the plant's size or age. We found that results using the original categories established by the authors were similar to those using the standardized categories (see Fig. \ref{fig:figA6}).  Thus in the main text, we report only results using the standardized categories.

@@

We tested an effect of plant stage on the relationship between growth rate and traits using a mixed-effect model, in which it was assumed that studies within a stage shared a common mean effect, but that there was also random variation among studies, in addition to sampling variation \citep{Zuur:2009cfa}. Inclusion of this random effect allowed us to account for heterogeneity among studies, for example in experimental type, and vegetation type. For each trait, we established whether stage could additional explain variance in $z$ by comparing models with a without a stage effect  using log-likelihood ratios \citep{Zuur:2009cfa}. If the log-likelihood test returned a significant p-value, the stage model fitted the data better than the model without. Estimates of the effect size for each stage were considered significant if the CIs did not overlap with zero.

@@

\clearpage

@@

% x <- remake::make("CleanData")
% unqiue(x$ref)
% nrow(x)
% x <- remake::make("CompleteData")
% unique(x$ref)
% nrow(x)
% x <- remake::make("IdealData")
% unique(x$ref)
% nrow(x)
\subsection*{Representation of stages among 19 traits}
In total, correlations had been reported for at least two different stages across 18 traits out of the 19 targeted (Fig. \ref{fig:fig1}). We had thought it likely that for some traits correlations would be reported only for one size class. But this limitation applied only for LAR and NARmass, where studies appeared confined to seedling stages. However, for a majority of traits the number of observations across stage was higthly unbalanced; for one-third of the 19 traits $>$ 75\% of correlations with growth reported come from a single stage (Fig. \ref{fig:fig1}).

@@

For Aarea, results were consistent with a positive correlation with RGR at seedling stage. However, no conclusion can be drawn about a change along plant stages; as only four correlation coefficients were reported for this trait, and three of these were measured at the seedling stage (Fig. \ref{fig:fig3}e).

@@

We now report results from the entire dataset to determine if additional data would alter or reinforce our conclusions, drawn sometimes with limited replicates. This entire dataset included correlations measured with both AGR and RGR (Fig. \ref{fig:fig4}), across at least two species (but measured using several individuals), and did not control for the source of trait or growth (i.e. trait could be measured on a different stage) or environmental conditions. The relaxed criteria considerably increased the number of replicates, from $n = 137$ to $n = 267$ correlations across 93 articles. However, the additional replicates were unequally distributed across trait-stage combinations. Some trait-stage combinations already with a lot of data got even more support (e.g. from 25 to 59 correlation coefficients reported for SLA at seedling stage), while others still had few replicates. On the 13 traits-stage combinations studied in total, six had $\le$ five correlation coefficients: WD-seedling, Hmax-sapling, Hmax-seedling, Seed mass-sapling, Aarea-adult and Aarea-sapling.

@@

Results from the entire dataset reinforced our conclusions for both SLA and WD, set appart alternative explanations for seed mass, and allowed to compare correlations with growth along plant stages for Aarea. For SLA and WD, the number of replicates doubled by using the entire dataset, and effect sizes and overall results were similar to those reported above using the ''ideal'' subset (Fig. \ref{fig:fig4})). As expected for seed mass, the correlation with growth at adult stage was significantly different from that observed at seedling stage (see CIs in Fig. \ref{fig:fig4}c, and model fit improved when stage was incorporated). For Aarea, the additional replicates in the entire dataset allowed us to compare the effect size across plant stage. The results were consistent with the prediction for a positive correlation with RGR across adult and seedling stages (Fig \ref{fig:fig4}e; CI not overlapping zero). Comparison of Aarea effect between the ``ideal'' subset (Fig. \ref{fig:fig3}d) and entire dataset (Fig. \ref{fig:fig4}d), indicated little effect of stress (here mainly low light) on the correlation with RGR at seedling stage; as the correlation was still positive when 8 correlations measured under low light conditions were added. For Hmax, there were still too few replicates to conclude undoubtely on the correlation to be absent for smaller sizes.

We had hypotheses for seed mass and SLA, about a change in the correlation between trait and growth as AGR was used instead of RGR, or as mass was used instead of dimater and height. For seed mass, effects size differed significantly between RGR and AGR at seedling stage; as expected the correlation with RGR was positive and shifted to non significant with AGR (Fig \ref{fig:fig4}d; CI not overlapping). For SLA, the correlation was expected to shift from positive to non significant if AGRmass was used instead of RGRmass for seedling stage; here none of the data have been measured with AGRmass at seedling stage (Fig. \ref{fig:figA8}). 
% Why a change with WD when it is measured in mass? All WD seedling are measured in RGRmass, and there are some mass data for both other stages too.  
For other traits, effect size were generally consistent between RGR and AGR when both measurement were available. Some differences occured but they were not easily attributed to any single cause. For instance for WD, the correlation at juvenile stage shifted from negative with RGR to non significant with AGR (i.e CI included zero, Fig. \ref{fig:fig4}b). The correlation coefficients recorded for AGR came from \citet{Augspurger:1984ct} and unlike others they used wood density values from adult plants together with growth from juveniles.

@@

In total, 103 articles published between 1983 and 2014, and containing 551 correlations linking growth rate to any of the 19 traits across a range of plant stages were identified. Data reported here were variable in multiple ways. Climate zones spanned from temperate regions to tropics (map in Fig. \ref{fig:figA2}). Different experiment types (laboratory, field, plantation or greenhouse), growth measurements and growth form (tree, woody or across growth form) were included. The number of species used for each correlation ranged from 2 to 300 species, with an average of 31. Given this variety, it was not surprising to find a wide range of $r$ reported for each trait. For instance, $r$ varied from -0.98 to 0.99 for SLA, and from -0.74 to 0.74 for WD (Fig. \ref{fig:figA3}). Growth form, experiment type and growth measurement type were not strong predictors in general (Fig. \ref{fig:figA6}); suggesting confounding with these factors is not a likely explanation of the major results reported.

>>>>

\input{ms-common-defs}
\date{}

\bibliographystyle{mee}
\mstype{Research Article}
\keywords{maximum height, photosynthetic rate, ontogenetic stage, growth strategy, seed mass, specific leaf area, wood density}

@@

\item We propose that traits should influence growth in a way that depends on the size of individual plants. We outline mechanisms and hypotheses based on new theoretical work, and test these predictions in tree species using a meta-analysis of 103 studies ($>$ 500 correlations) for five traits (specific leaf area, wood density, maximum height, seed mass and maximum assimilation rate). We also recorded data for 14 others traits commonly used in the trait literature. To capture the effects of plant size, we tested for a shift in the direction of correlation between growth rates and traits across three ontogenetic stages: seedling, sapling and adult.

\item Results were consistent with predictions, even with some limitations arising from unequal numbers of observation across ontogenetic stages. Specific leaf area was correlated with relative growth rate (RGR) in seedlings but not in adult plants. Correlations of growth with wood density and assimilation rate were not affected by stage. Seed mass and maximum height were correlated with RGR only in one ontogenetic stage category: seedlings and adults, respectively.

\item Although we were able to confirm several of our theoretical predictions, major knowledge gaps size still exist, despite a sizeable literature on traits and growth. For example, for one-third of the trait considered, the majority ($>$ 75\%) of reported correlations with growth came from a single ontogenetic stage.

\item Synthesis: We show how trait-growth correlations change in a predictable way with plant size. Our understanding of plant strategies should shift away from attributing species as having slow or fast growth throughout their life, in favour of attributing growth trajectories.
\end{enumerate}

@@

Meanwhile the possibility has been raised that trait-growth correlations may change systematically with size \citep{King-1999, Falster:2011ii, Ruger:2012jv, Iida:2014ep, Iida:2014hq}. Size has long been known to play a central role in organism design and function \citep{Vogel:1988ux, Vogel:2003wb}, including that of trees \citep{Farnsworth:1995im, Givnish:1995ta, King-2011}. As plants grow larger, a greater amount of the energy is devoted towards building and maintaining non-photosynthetic tissues \citep{Givnish:1995ta,King-2011}. A recent model by \citet{Falster:2011ii} suggests this change shifts the benefits and costs of some trait-based trade-offs, such that the net effect of the trait on growth changes with size. Thus the absence of a correlation between SLA and growth at larger sizes may in fact be expected, when potential influences of traits on plant function are considered \citep{Falster:2011ii}. While recent empirical studies support the idea that the relationship between traits and growth rates may change with size (\citealt{Iida:2014ep, Iida:2014hq}), the trend and mechanisms behind this variation remain unclear.

In this paper, we investigate for five prominent traits (specific leaf area = SLA, wood or stem tissue density = WD, maximum leaf photosynthetic capacity per unit area = Aarea, asymptotic or maximum height = Hmax, and seed mass):
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions derived from a recently developed mechanistic model \citep{Falster:2011ii} (see ``Theoretical expectations'')?
\end{enumerate}
To address these questions we used a meta-analytic approach, compiling data from diverse studies reporting correlation coefficients between traits and growth. While most studies investigating the link between traits and growth have not primarily been concerned with the effects of plant size, they have generally reported trait-growth correlations for identifiable ontogenetic stages (i.e. seedling, sapling, and adult). We therefore used ontogenetic stage as a surrogate for plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.
We chose to assess our hypotheses via a meta-analysis because there already exist a considerable number of studies with salient information. Using these studies, we were able to assess our hypotheses across a much broader number of species and vegetation types than would have been possible in an experiential study. While, we anticipated a reasonably-sizeable literature on trait and growth to be available, we were uncertain about the representation of all trait-stages combinations. An additional goal of our study was therefore to identify gaps in the knowledge needed to improve our understanding of the link between trait and growth.

@@

Based on the above model and this single assumption about growth, we outline five hypotheses that arise by considering the influence of trait variation on elements of equations  \ref{eq:eq0}-\ref{eq:eq4} and how these may interact with plant size (see Table \ref{tab:trade-offs}). These hypotheses reflect a direct mechanistic causation between traits and growth rate. It is possible that other correlations may arise indirectly. Nevertheless, for purposes of this paper we adopt as working hypotheses only those arising from direct mechanistic causation. Moreover, we consider how predictions may vary when expressed on absolute and relative terms, for each of height, diameter, and mass-based growth (see Table \ref{tab:trade-offs}). For many trait and growth combinations, the predicted effect of the trait on growth is similar when expressed on relative and absolute terms. However, this is not true for all traits (details given below). We therefore adopted a two-stage approach to our analysis, whereby we first assess patterns based on measurements of relative growth rate.

@@

Plants with high SLA can deploy a unit of leaf area more cheaply. However, high-SLA species also have shorter leaf lifespan, and must use more energy replacing leaves lost via turnover \citep{Wright:2004jb}. These two effects occur within the terms  $\frac{\ud B}{\ud t}$ and $\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$ in the above model. Whether plant growth rate increases with SLA depends on how the relative magnitude of the benefit (cheap leaf construction, via $\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$ ) and cost (higher leaf turnover, via $\frac{\ud B}{\ud t}$) of increased SLA change with size (Table \ref{tab:trade-offs}). 
During early life, when leaf comprises a large part of the plant, decreasing SLA has an overwhelmingly positive effect on height and diameter growth rate, because the effect of reducing $\frac{\ud M_{\rm l}}{\ud A_{\rm l}}$ has a large effect compared to the other terms in eq. \ref{eq:daldmt}. As plants increase in size, however, the cost of building sapwood ($\frac{\ud M_{\rm s}}{\ud A_{\rm l}}$) increases disproportionately, decreasing the benefit of cheaper leaf construction in eq.  \ref{eq:daldmt}. Consequently, the relative effect of leaf turnover comes to dominate at larger sizes, and as such, the effect of SLA on height and diameter growth shifts from positive to negative as plants increase in size. A similar shift is expected for growth in mass when expressed in relative terms. However, when expressed as absolute growth, the predicted relationship is negative. This occurs because of the increased turnover of high SLA species.

@@

Cheaper stem construction (low WD or stem density) increases the marginal rate of leaf deployment per total mass increment ($\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$), and consequently increases growth rate in height and diameter (eqs. \ref{eq:eq2}-\ref{eq:eq4}; Table \ref{tab:trade-offs}). Disadvantages associated with cheap stem construction take the form of higher mortality and so are not reflected in growth equations. Under this scenario, wood density is negatively correlated to growth rate across all ontogenetic stages. Because the relative expenditure on stem compared to leaf increases as plants grow larger, the relationship between WD and growth may also strengthen in larger plants. A similar shift is expected for growth in mass when expressed in relative terms. However, when expressed as absolute growth, the predicted relationship is non-significant \citep[see also][]{Enquist-1999}.

@@

As plants approach maximum height they increasingly divert energy away from growth and into reproduction (see \citealt{Thomas:1996do,Thomas-2011, Wenk:2015jz}). This effect occurs in the term $\frac{\ud M_{\rm t}}{\ud B}$ in the above model eqs. (\ref{eq:eq2}-\ref{eq:eq4}). For seedlings, where all individuals are investing solely in growth, we expect no influence of Hmax on growth (i.e. $\frac{\ud M_{\rm t}}{\ud B} =1$). But once some species begin to produce seed, then species closer to their Hmax value are expected to be allocating more to seed production and less to vegetative growth. Hence when comparing across plants at a given size, species with smaller Hmax are expected to show slower vegetative growth rates. Similar correlations are expected no matter what measure of  growth is used (Table \ref{tab:trade-offs}).

@@

The main effect of increased seed size is to produce a larger overall seedling. Under similar light conditions, larger seedlings are therefore predicted to have larger absolute growth rates. However, because the ratio of leaf area to total biomass decreases with plant size, larger seedlings are predicted to have lower mass-based growth rates during the earliest phases of growth \citep{Turnbull:2012ew}. Yet, as plants grow, we expect the effect of this initial size difference to be minimal, and the correlation between seed size and growth rate should therfore disappear. Thus for larger plants, seed mass is not expected to have any distinct influence on growth rates  (Table \ref{tab:trade-offs}).

@@

We expect greater photosynthetic capacity to translate into faster growth rates by increasing the rate of biomass production (Table \ref{tab:trade-offs}), at least at high light where Aarea is typically measured. This effects occurs within the term  $\frac{\ud B}{\ud t} $ in the above model. The effects of benefits (higher assimilation) and costs (higher respiration) of increased Area occur within the same element of eq. \ref{eq:eq0}, and because assimilation is typically higher than respiration in high light, the net effect will be positive (Table \ref{tab:trade-offs}). It is uncertain how much this effect might translate to faster growth in low light.

@@

\subsubsection*{Ontogenetic stage categories}

We used categorical groupings by ontogenetic ``stage'' to test our hypotheses regarding the effects of traits on growth rates. The authors of the 103 articles used in this analysis reported their plants as belonging to any of five different stages: ``seedling'', ``juvenile'', ``sapling'', ``adult'', ``mixed''. While these categories have been established according to the expertise of the authors, we were concerned about the risk of assigning plants of similar size to different categories when compiling data across multiple studies and vegetation types. This risk turned out to be small: overall, we found that the original thresholds in size/age used by the authors to distinguish different stages were relatively compatible (Fig. \ref{fig:fig2}). In particular, there was a clear break between seedlings and saplings in both age (at 1 yr) and height (at 0.5 m). There was also a clear break between saplings and adults in diameter (at 10 cm). Comparing these values to the bivariate relationships observed between variables (age, height, diameter) in the ``Biomass And Allometry Database'' \citep{Falster:2015}, demonstrated that the thresholds identified above for each stage transition also aligned relatively well with one another (Fig. \ref{fig:figA1}). Thus we adopted the above values as cut-offs for defining stage boundaries.

Using the identified boundaries among stages we were able to assign all studies into one of three classes: ``seedling'', ``sapling'', or ``adult''. The group ``juvenile'' was largely overlapping with the three others, and so we merged this category with one or the other on a case by case basis. Using the above cut-offs, we were also able to assign a stage category to studies for which no stage was distinctly mentioned by the authors (5\% of the articles, noted as ``mix'' in Fig. \ref{fig:fig2}). Only 7 of 103 articles were ambiguous on their ontogenetic stage attribution. Half of these equivocal cases involved the juvenile category. We attributed a study to a new stage category only when 95\% of the recorded range of sizes for that study lay within a particular stage category. Other equivocal cases were caused by ambiguity in the usage of ``adult'' by the authors. Some plants with a large stature were not described as adult by authors because they were still non reproductive \citep{King:2006he}. Conversely plants with small stature -- because stunted by wind for example \citep{Stratton:2001ck} -- were sometimes assessed as adult when reproductive. For purposes of the present paper, we attributed a study to the adult stage when the authors identified the plants as reproductive, and to sapling stage when they identified the plants as non-reproductive, regardless of the plant's size or age. We found that results using the original categories established by the authors were similar to those using the standardized categories (see Fig. \ref{fig:figA6}).  Thus in the main text, we report only results using the standardized categories.

@@

We tested an effect of ontogenetic stage on the relationship between growth rate and traits using a mixed-effect model, in which it was assumed that studies within a stage shared a common mean effect, but that there was also random variation among studies, in addition to sampling variation \citep{Zuur:2009cfa}. Inclusion of this random effect allowed us to account for heterogeneity among studies, for example in experimental type, and vegetation type. For each trait, we established whether stage could additional explain variance in $z$ by comparing models with a without a stage effect  using log-likelihood ratios \citep{Zuur:2009cfa}. If the log-likelihood test returned a significant p-value, the stage model fitted the data better than the model without. Estimates of the effect size for each stage were considered significant if the CIs did not overlap with zero.

@@

@@

% x <- remake::make("CleanData")
% unqiue(x$ref)
% nrow(x)
% x <- remake::make("CompleteData")
% unique(x$ref)
% nrow(x)
% x <- remake::make("IdealData")
% unique(x$ref)
% nrow(x)
\subsection*{Representation of stages among 19 traits}
In total, correlations had been reported for at least two different stages across 18 traits out of the 19 targeted (Fig. \ref{fig:fig1}). We had thought it likely that for some traits, correlations would be reported only for one growth stage. But this limitation applied only for LAR and NARmass, where studies appeared confined to seedling stages. However, for a majority of traits the number of observations across stage was highly unbalanced; for one-third of the 19 traits $>$ 75\% of correlations with growth reported come from a single stage (Fig. \ref{fig:fig1}).

@@

For Aarea, results were consistent with a positive correlation with RGR at seedling stage. However, no conclusion can be drawn about a change along ontogenetic stages; as only four correlation coefficients were reported for this trait, and three of these were measured at the seedling stage (Fig. \ref{fig:fig3}e).

@@

We now report results from the entire dataset. Our aim is to determine if additional data alters or reinforce the above conclusions, by increasing replication. This entire dataset included correlations measured with both AGR and RGR (Fig. \ref{fig:fig4}), across at least two species (but measured using several individuals), and did not control for the source of trait or growth (i.e. trait could be measured on a different ontogenetic stage to growth data) or environmental conditions. The relaxed criteria considerably increased the number of replicates, from $n = 137$ to $n = 267$ correlations across 93 articles. However, the additional replicates were unequally distributed across trait-stage combinations. Some trait-stage combinations already with a lot of data got even more support (e.g. from 25 to 59 correlation coefficients reported for SLA at seedling stage), while others still had few replicates. On the 13 traits-stage combinations studied in total, six had $\le$ five correlation coefficients: WD-seedling, Hmax-sapling, Hmax-seedling, Seed mass-sapling, Aarea-adult and Aarea-sapling.

@@

Results from the entire dataset reinforced our conclusions for both SLA and WD, set apart alternative explanations for seed mass, and allowed us to compare correlations with growth along ontogenetic stages for Aarea. For SLA and WD, the number of replicates doubled by using the entire dataset, and effect sizes and overall results were similar to those reported above using the ''ideal'' subset (Fig. \ref{fig:fig4})). As expected for seed mass, the correlation with growth at adult stage was significantly different from that observed at seedling stage (see CIs in Fig. \ref{fig:fig4}c, and model fit improved when stage was incorporated). For Aarea, the additional replicates in the entire dataset allowed us to compare the effect size across ontogenetic stage. The results were consistent with the prediction for a positive correlation with RGR across adult and seedling stages (Fig \ref{fig:fig4}e; CI not overlapping zero). Comparison of Aarea effect between the ``ideal'' subset (Fig. \ref{fig:fig3}d) and entire dataset (Fig. \ref{fig:fig4}d), indicated little effect of stress (here mainly low light) on the correlation with RGR at seedling stage; as the correlation was still positive when 8 correlations measured under low light conditions were added. For Hmax, there were still too few replicates to conclude undoubtely on the correlation to be absent for smaller sizes.

We had hypotheses for seed mass and SLA, about a change in the correlation between trait and growth as AGR was used instead of RGR, or as mass was used instead of dimater and height. For seed mass, effects size differed significantly between RGR and AGR at seedling stage; as expected the correlation with RGR was positive and shifted to non significant with AGR (Fig \ref{fig:fig4}d; CI not overlapping). For SLA, the correlation was expected to shift from positive to non significant if AGRmass was used instead of RGRmass for seedling stage; here none of the data have been measured with AGRmass at seedling stage (Fig. \ref{fig:figA8}). 
% Why a change with WD when it is measured in mass? All WD seedling are measured in RGRmass, and there are some mass data for both other stages too.  
For other traits, effect size were generally consistent between RGR and AGR when both measurement were available. Some differences occurred but they were not easily attributed to any single cause. For instance for WD, the correlation at juvenile stage shifted from negative with RGR to non significant with AGR (i.e CI included zero, Fig. \ref{fig:fig4}b). The correlation coefficients recorded for AGR came from \citet{Augspurger:1984ct} and unlike others they used wood density values from adult plants together with growth from juveniles.

@@

In total, 103 articles published between 1983 and 2014, and containing 551 correlations linking growth rate to any of the 19 traits across a range of ontogenetic stages were identified. Data reported here were variable in multiple ways. Climate zones spanned from temperate regions to tropics (map in Fig. \ref{fig:figA2}). Different experiment types (laboratory, field, plantation or greenhouse), growth measurements and growth form (tree, woody or across growth form) were included. The number of species used for each correlation ranged from 2 to 300 species, with an average of 31. Given this variety, it was not surprising to find a wide range of $r$ reported for each trait. For instance, $r$ varied from -0.98 to 0.99 for SLA, and from -0.74 to 0.74 for WD (Fig. \ref{fig:figA3}). Growth form, experiment type and growth measurement type were not strong predictors in general (Fig. \ref{fig:figA6}); suggesting confounding with these factors is not a likely explanation of the major results reported.

Overall our study met quality criteria expected for meta-analysis in ecology (Table \ref{tab:meta}). The impact of publication bias was seemingly trivial (See Appendix \ref{app:growth-equation}, Fig. \ref{fig:figA7}). The extent of heterogeneity was high, but was substantially explained by our core hypothesis of an effect of ontogenetic stage for SLA and seed mass (Appendix \ref{app:supp_info_analyses}). There was no indication of a temporal change in the effect size being reported within stages (Appendix \ref{app:supp_info_analyses}, Fig. \ref{fig:figA9}). Finally, our results were unlikely to be biased by dominance of one or a few studies that measured a large number of responses, since when multiple responses were condensed into a single average within a trait and within a study, outcomes were very similar (Fig. \ref{fig:figA5}).
